---
title: "Sawfish"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(ggdist)
library(tidybayes)
```


## Supplementing length data

By using the rostrum length to infer the total length we could increase the sample size by 6. It's not a dramatic increase, but given the paucity of data it seems to be a valuable exercise. Details of those observations are shown below. 

```{r}
drake::loadd(encounters)
encounters %>%
  count(is.na(rostrum_length_cm), is.na(total_length_cm), is.na(reported_collection_year))

encounters %>%
  filter(is.na(total_length_cm), !is.na(rostrum_length_cm), !is.na(reported_collection_year)) %>%
  select(country, species, ocean, sex, source_type)
```

## Allometric relationship

We use another model to the allometric relationship between rostrum and total length. This allows us to estimate the total length for those encounters that only have rostrum length. 
As expected the uncertainty is larger when the species is not known. Note that this would be greatly improved if we could source existing data from the literature. 

```{r pressure, echo=FALSE}
drake::loadd(allometric_model)
drake::loadd(allometric_model_data)

add_fitted_draws(
  newdata = tidyr::expand_grid(
    rostrum_length_cm = seq(min(allometric_model$data$rostrum_length_cm),
                          max(allometric_model$data$rostrum_length_cm), 
                          length.out = 10), 
    species = unique(allometric_model$data$species)), 
  model = allometric_model) %>%
  ungroup() %>%
  mutate(across(c(rostrum_length_cm, .value), exp)) %>% 
  ggplot(aes(x = rostrum_length_cm/100, y = .value/100)) +
  stat_lineribbon(aes(alpha = forcats::fct_rev(ordered(stat(.width)))), 
                  .width = c(0.05, 0.66, 0.95), size = 0.5, fill = "grey30") +
  geom_point(data = allometric_model_data, 
             aes(y = exp(total_length_cm)/100, 
                 x = exp(rostrum_length_cm)/100)) + 
  facet_wrap("species") +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "Rostrum length (m)", 
       y = "Total length (m)")

```

A closer inspection to the allometric coefficient reveals that the rostrum most likely grows isometrically with total length. 

```{r fig.height=3}
coefficients <- spread_draws(allometric_model, b_Intercept, b_rostrum_length_cm) %>%
  mutate(alpha = 1 / b_rostrum_length_cm, 
         log_b = b_Intercept / b_rostrum_length_cm * (-1)) %>%
  select(-b_Intercept, -b_rostrum_length_cm) 


coefficients %>%
  ggplot(aes(x = alpha)) +
  stat_halfeye() +
  theme_minimal() + 
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank()) + 
  labs(x = "Allometric coefficient (⍺)")
```

## Length data

Let's explore the data using different strata. This will allow us to better identify how data is distributed across different categories and which ones have the potential to introduce biases in the data. 

Most available length-year data is from *Pristis pristis* in the Panama's Pacific. 
There is also a strong bias of the year of different source types. It would be important to take these biases into account. 

In addition a group of records from museums that have small individuals (< 1.5 meters) that might not be easily explained by the current strata. Similar cases for literature sourced encounters where a range of sizes is available for a given year. 

```{r, fig.height=12}
drake::loadd(length_model_data)

bp <-  length_model_data %>%
  filter(!is.na(reported_collection_year), !is.na(total_length_cm)) %>%
  ggplot(aes(x = reported_collection_year, y = total_length_cm)) +
  geom_linerange(aes(ymin = total_length_cm - total_length_cm_se, 
                     ymax = total_length_cm + total_length_cm_se)) +
  theme_minimal() +
  theme(axis.title.x = element_blank()) +
  labs(y = "total length (cm)")

p1 <- bp + geom_point(aes(colour = country), shape = 21) 
p2 <- bp + geom_point(aes(colour = species), shape = 21)
p3 <- bp + geom_point(aes(colour = ocean), shape = 21)
p4 <- bp + geom_point(aes(colour = sex), shape = 21)
p5 <- bp + geom_point(aes(colour = source_type), shape = 21)

p1 + p2 + p3 + p4 + p5 + plot_layout(ncol = 1)
```

## Model results

```{r, fig.height=6, fig.width=10}
drake::readd(fig_max_length_vs_time) +
  plot_annotation(title = "Modelling max total length")

drake::readd(fig_length_vs_time) +
  plot_annotation(title = "Modelling all total length")
```

## Encounters

```{r}
drake::loadd(encounters)

enc_data <- encounters %>%
  filter(reported_collection_year > 1800) %>%
  mutate(source_type = fct_reorder(source_type, reported_collection_year, median), 
         country = fct_infreq(country))

enc_data %>%
  ggplot(aes(x = reported_collection_year, fill = source_type)) + 
  geom_histogram(binwidth = 10, center = 2000) +
  facet_wrap("country", ncol = 2) +
  # scale_fill_brewer(palette = "Set1") + 
  # scale_y_continuous(sec.axis = sec_axis(~ . / nrow(enc_data), 
                                         # labels = scales::label_percent(), name = "Percentage of encounters")) +
  scale_fill_viridis_d() +
  # scale_x_binned(stat  = "co") +
  theme_minimal() +
  theme(legend.position = "top", 
        legend.title = element_blank()) +
  labs(x = "Reported collection year", 
       y = "Number of encounters")
```

```{r}
library(rnaturalearth)
library(rnaturalearthhires)
world <- ne_countries(country = c("Colombia", "Panama", "Costa Rica"), scale = "large", returnclass = "sf")

map_data <-
  encounters %>%
    filter(reported_collection_year > 1800) %>%
  mutate(source_type = fct_reorder(source_type, reported_collection_year, median), 
         across(c(longitude, latitude), as.numeric)) %>%
  filter(latitude > 0)

p1 <- ggplot(data = map_data) + 
  geom_sf(data = world, inherit.aes = F, alpha = 0.5, colour = "grey30") +
  geom_point(data = map_data, aes(x = longitude, y = latitude, fill = source_type), shape = 21) + 
  scale_fill_viridis_d() +
  theme_minimal() +
  theme(axis.title = element_blank()) +
  coord_sf(xlim = c(-83.9, -73), ylim = c(4.1, 11.9), expand = FALSE) +
  scale_x_continuous(breaks = seq(-90, 90, by = 2)) +
  scale_y_continuous(breaks = seq(-90, 90, by = 2)) +
  labs(fill = "Source")

p2 <- ggplot(data = map_data) + 
  geom_sf(data = world, inherit.aes = F, alpha = 0.5, colour = "grey30") +
  geom_point(data = map_data, aes(x = longitude, y = latitude, fill = reported_collection_year), shape = 21) + 
  scale_fill_stepsn(breaks = seq(1000, 3000, by = 25), colours = viridis::viridis(2)) +
  theme_minimal() +
  theme(axis.title = element_blank()) +
  coord_sf(xlim = c(-83.9, -73), ylim = c(4.1, 11.9), expand = FALSE) +
  scale_x_continuous(breaks = seq(-90, 90, by = 2)) +
  scale_y_continuous(breaks = seq(-90, 90, by = 2)) +
  labs(fill = "Year")

p1 + p2 + plot_layout(ncol = 1) 
```

