---
title: "Paper"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
library(brms)
library(tidybayes)
library(broom.mixed)
ci_width <- 0.9
```

## Methods

A (bayesian) regression model was constructed to investigate how sawfish size has changed over time. 
Records in which both total length and date were available were selected. 
These records were supplemented with estimations of total length was estimated using based on rostrum length and the allometric relationship between this body part and the total length (Supporting Information S1). 
This analysis was focused in the *Pristis pristis* size, as this was the species where most complete records were available. 
These records were further filtered to include only the largest total length measurement for a given year/source type combination. 
This was done so, because due to the limited number of records, it is unlikely the data provides a complete picture of the size distribution changes. 
Many of the records are likely to correspond to neonate and juvenile individuals and therefore may offer little information about overal size changes. 
Furthermore, by focusing in the largest individuals, across time and a change on the maximum observed size provides information about ... [Juli: please add here an explanation of why max size is also a biologically relevant metric].
Importantly, the results do not qualitatively change whether we include all records or just those with the largest individuals (Supporting Information S2).

The response variable in the model was total length and the explanatory variable was the year of record. 
The year of record was scaled to have a zero mean and unit variance prior to inclusion in the model. 
In addition, to control for potential biases in the data random intercepts for the country of record and the source of the record were calculated (museum collection, literature, ISED, interview and non-interview observation). 
Total length was modelled using a Binomial distribution. 
As opposed to a Poission distribution, which is also commonly used to model positive-only length data, as this distribution is ideal for modelling integer values within a specified range; this is important as total length has an upper bound. 
This upper bound was set at 120% the length of the largest measured specimen. 
To account for overdispersion in the data each observation was included as an additional random intercept. 
A weakly informative prior based on a Student-t distribution with 3 degrees of freedom, a median of zero, and a scale of two, was used for all model coefficients. 
The bayesian model was fitted using `brms 2.14.4` (B端rkner 2017, 2018) which uses a Hamiltonian Markov Chain Monte Carlo algorithm implemented in `Stan` (Stan Development Team, 2020). 
The model was fitted using four chains of 5000 iterations each (including 2500 for warm-up).

## Results

```{r results-allometry, include = FALSE}
drake::loadd(allometric_model)
allometric_coefficients <- spread_draws(allometric_model, r_species[species,parmeter], b_Intercept, b_rostrum_length_cm) %>%
  mutate(species_intercept = b_Intercept + r_species, 
         alpha = 1 / b_rostrum_length_cm, 
         log_b = species_intercept / b_rostrum_length_cm * (-1)) %>%
  select(-b_Intercept, -b_rostrum_length_cm) %>%
  filter(species == "Pristis.pristis") %>%
  median_qi(alpha, log_b, .width = ci_width) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

```

```{r results-max-length, include = FALSE}
drake::loadd(max_length_model_data)
drake::loadd(max_length_model)
drake::loadd(fig_max_length_vs_time)

estimated_lengths <- max_length_model_data %>%
  filter(total_length_cm_se > 0)

fitted_draws_max_length_model <- attr(fig_max_length_vs_time, "data") %>%
  group_by(.draw, source_type) %>%
  mutate(ten_year_diff = (.value - lead(.value, 10))/.value) %>%
  filter(!is.na(ten_year_diff), 
         source_type == "Overall") %>%
  group_by(source_type) 

scaled_year_par <- max_length_model %>%
  spread_draws(b_scaled_year)

pd_year <- scaled_year_par %>%
  summarise(p = sum(b_scaled_year < 0)/n())

est_year <- median_qi(scaled_year_par, .width = ci_width) %>%
  mutate(across(where(is.numeric), scales::number_format(0.01)))

decrease_per_decade <- c(0.66, 0.9, 0.95) %>%
  map_dfr(~median_qi(fitted_draws_max_length_model, ten_year_diff, .width = .)) %>%
  mutate(across(c(ten_year_diff, .lower, .upper), scales::percent_format())) %>%
  split(list(.$.width)) %>%
  purrr::array_tree()
```   

The relationship between rostrum and total length for the sawfish *Pristis pristis* was more or less isometric (Supporting Information S1). 
Indeed the estimated allometric coefficient between these two metric was `r allometric_coefficients$alpha` (`r scales::percent(ci_width)` credible interval `r allometric_coefficients$alpha.lower` to `r allometric_coefficients$alpha.upper`). 

The total length model was based on `r nrow(max_length_model_data)` records. 
In `r nrow(estimated_lengths)` of them, total length was estimated using the allometric relationship between rostrum and total length.
There is a `r scales::percent(pd_year$p, accuracy = 0.1)` probability that the total length of the largest observed sawfish individuals (*Pristis pristis*) has decreased between `r min(max_length_model_data$reported_collection_year)` and `r max(max_length_model_data$reported_collection_year)` (Figure XXa, Table S2). 
The median rate of this decrease is `r decrease_per_decade[[2]]$ten_year_diff` per decade (`r scales::percent(ci_width)` CI `r decrease_per_decade[[2]]$.lower` to `r decrease_per_decade[[2]]$.upper`; Figure XXb).
We found no substantial differences between the intercepts of different record sources (Figure S2).

```{r length-resuts-figure, echo = F, fig.width=4, fig.height=4}
fig_max_length_vs_time +
  plot_annotation(tag_levels = "a", tag_suffix = ")")
```

*Figure XX: Maximum length of sawfish Pristis pristis has very likely declined over time. (a) Dots indicate length of records included in the model; vertical bars indicate the standard error of total lengths estimated from rostrum length. Grey bands indicate 66 and 90% credible intervals. (b) Distribution of the rate of decrease in total length per decade. The point depict the median value and lines correspond to 66 and 90% credible intervals.*

## Supporting information

### S1: Estimating total length from rostrum length

```{r results-supp-allometry, include = F}
r2 <- bayes_R2(allometric_model, probs = c((1-ci_width)/2, 1 - (1-ci_width)/2)) %>%
  round(2)
```

A (bayesian) regression model was constructed to investigate the allometric relationship between rostrum and total length. 
A total of `r nrow(allometric_model$data)` records in which both of these measurements were available were selected. 
As is common in allometric estimation, both of these metrics were log transformed.
Then a linear regression was fitted; total length was the response variable and rostrum length the explanatory variable. 
This was done because our primary objective is the estimation of total length and not the estimation of the allometric coefficient (in which case a major axis regression would have been more appropriate).
In addition, a random intercept was calculated for each species type. 
In the absence of previous information in the literature, we used weakly informative prior based on a Student-t distribution with 3 degrees of freedom, a median of zero, and a scale of two, was used for all model coefficients except the one corresponding to the slope of the rostrum-total length relationship where a similar Student-t distribution with a median of one was used. 
The bayesian model was fitted using brms 2.14.4 (B端rkner 2017, 2018) which uses a Hamiltonian Markov Chain Monte Carlo algorithm implemented in Stan (Stan Development Team, 2020). 
Specifically, the model was fitted using four chains of 5000 iterations each (including 2500 for warm-up). 

The model accounted for a large proportion of variance in total length (R^2^ = `r r2[[1]]`, 90% CI `r r2[[3]]` to `r r2[[4]]`). 
Total length scales isometrically with rostrum length in *Pristis pristis* (Figure S1; Table S1). 
Using the model coefficients, the estimated allometric coefficient between these two metric was `r allometric_coefficients$alpha` (`r scales::percent(ci_width)` credible interval `r allometric_coefficients$alpha.lower` to `r allometric_coefficients$alpha.upper`). 
The estimated allometric intercept was `r allometric_coefficients$log_b` (`r scales::percent(ci_width)` CI `r allometric_coefficients$log_b.lower` to `r allometric_coefficients$log_b.upper`).

### S2: Including all size records

```{r results-all-length, include = FALSE}
drake::loadd(length_model_data)
drake::loadd(length_model)
drake::loadd(fig_length_vs_time)

estimated_lengths_all <- length_model_data %>%
  filter(total_length_cm_se > 0)

fitted_draws_length_model <- attr(fig_length_vs_time, "data") %>%
  group_by(.draw, source_type) %>%
  mutate(ten_year_diff = (.value - lead(.value, 10))/.value) %>%
  filter(!is.na(ten_year_diff), 
         source_type == "Overall") %>%
  group_by(source_type) 

scaled_year_par_all <- length_model %>%
  spread_draws(b_scaled_year)

pd_year_all <- scaled_year_par_all %>%
  summarise(p = sum(b_scaled_year < 0)/n())

est_year_all <- median_qi(scaled_year_par_all, .width = ci_width) %>%
  mutate(across(where(is.numeric), scales::number_format(0.01)))

decrease_per_decade_all <- c(0.66, 0.9, 0.95) %>%
  map_dfr(~median_qi(fitted_draws_length_model, ten_year_diff, .width = .)) %>%
  mutate(across(c(ten_year_diff, .lower, .upper), scales::percent_format())) %>%
  split(list(.$.width)) %>%
  purrr::array_tree()
```   

The results are qualitatively similar when all observations in which total length is available, not just those for the largest individuals per year, are included in the model. 
The total length model with all records was based on `r nrow(length_model_data)` records. 
In `r nrow(estimated_lengths_all)` of them, total length was estimated using the allometric relationship between rostrum and total length.
In this alternative model, total length of encountered sawfish (*Pristis pristis*) individuals was also very likely to decrease over the years (`r scales::percent(pd_year_all$p, accuracy = 0.1)` probability). 
The estimated coefficient for the year parameter was `r est_year_all$b_scaled_year` (`r scales::percent(ci_width)` CI `r est_year_all$.lower` to `r est_year_all$.upper`).
This coefficient corresponds to a median rate of decrease of `r decrease_per_decade_all[[2]]$ten_year_diff` per decade (`r scales::percent(ci_width)` CI `r decrease_per_decade_all[[2]]$.lower` to `r decrease_per_decade_all[[2]]$.upper`).

### Supplementary Figures

```{r rostrum-vs-total, echo = F, fig.height=3, fig.width=8}
drake::loadd(allometric_model_data)

add_fitted_draws(
  newdata = tidyr::expand_grid(
    rostrum_length_cm = seq(min(allometric_model$data$rostrum_length_cm),
                          max(allometric_model$data$rostrum_length_cm), 
                          length.out = 10), 
    species = unique(allometric_model$data$species)), 
  model = allometric_model) %>%
  ungroup() %>%
  mutate(across(c(rostrum_length_cm, .value), exp)) %>% 
  ggplot(aes(x = rostrum_length_cm/100, y = .value/100)) +
  stat_lineribbon(aes(alpha = forcats::fct_rev(ordered(stat(.width)))), 
                  .width = c(0.05, 0.66, 0.95), size = 0.5, fill = "black") +
  geom_point(data = allometric_model_data, 
             aes(y = exp(total_length_cm)/100, 
                 x = exp(rostrum_length_cm)/100), 
             shape = 21) + 
  facet_wrap("species") +
  theme_minimal() +
  theme(strip.text = element_text(face = "italic")) +
  theme(legend.position = "none") +
  labs(x = "Rostrum length (m)", 
       y = "Total length (m)")
```

*Fig S1: Rostrum and total length in sawfish scales isometrically. The panels show the relationship between rostrum and total length. Dots indicate observed data and grey bands indicate 66 and 90% credible intervals.*

```{r length-random, echo = F, fig.height=2, fig.width=4}
max_length_model %>%
  gather_draws(r_source_type[source_type,Intercept]) %>%
  ungroup() %>%
  mutate(source_type = snakecase::to_sentence_case(source_type), 
         source_type = fct_reorder(source_type, .value, .desc = T)) %>%
  ggplot(aes(x = .value, y = source_type)) +
  stat_pointinterval() +
  geom_vline(xintercept = 0, linetype = 2, size = 0.25) +
  theme_minimal() +
  theme(axis.title.y = element_blank(), 
        panel.grid.major.y = element_blank()) +
  labs(x = "Intercept")
```

*Figure S2: There were no substantial differences in the mean total length across different record sources. Points depict the estimated median intercept for each of the sources. Lines correspond to 66 and 90% credible intervals of the intercept.*

### Supplementary Tables

*Table S1: Model coefficients for the allometry model. Intervals correspond to 95% credible intervals*.
```{r allometric-model-table, warning = FALSE, echo = F}
allometric_model %>%
  broom::tidy() %>%
  select(-component, -std.error) %>%
  mutate_if(is.character, ~replace_na(., replace = "-")) %>%
  knitr::kable(digits = 2)
```

*Table S2: Model coefficients for the total length model. Intervals correspond to 95% credible intervals*.
```{r length-model-table, warning = FALSE, echo = F}
max_length_model %>%
  broom::tidy() %>%
  select(-component, -std.error) %>%
  mutate_if(is.character, ~replace_na(., replace = "-")) %>%
  knitr::kable(digits = 2)
```

## References

Paul-Christian B端rkner (2017). brms: An R Package for Bayesian Multilevel Models Using Stan. Journal of Statistical Software, 80(1), 1-28. doi:10.18637/jss.v080.i01

Paul-Christian B端rkner (2018). Advanced Bayesian Multilevel Modeling with the R Package brms. The R Journal, 10(1), 395-411. doi:10.32614/RJ-2018-017

Stan Development Team. 2020. Stan Modeling Language Users Guide and Reference Manual, 2.21.0 https://mc-stan.org
